Determining the alternatives for scalar implicature

### 2016-02-02 Paper/Paper.Rmd
second to last pass fixed x- alts labeling and some wording

In our current work, we use a computational framework to investigate the set of alternatives that best allow the model to predict human pragmatic judgments. We begin by presenting the computational framework (RSA) we use throughout the paper. We next describe a series of experiments designed to measure both the literal semantics of a set of scalar items (used to simulate different alternative sets available to listeners) and comprehenders' pragmatic judgments for these same items (used for model comparison). These experiments allow us to compare the effects of different alternative sets on our ability to model listeners' pragmatic judgments. To preview our results: we find that standard entailment alternatives do not allow us to fit participants' judgments, but that expanding the range of alternatives empirically (by asking participants to generate alternative messages) allows us to predict listeners' pragmatic judgments with high accuracy.

@@

```{r allScalesTable, fig.pos = "t", fig.width=3.5, fig.height=2.5, fig.cap = "Alternative sets used in Experiments 2 and 3. Green coloring denotes classic entailment items used in conditions 2a and 3a. Blue coloring denotes top-two alternatives added to entailment scales (from Experiment 1) in conditions 2b and 3b. Red coloring denotes ``neutral\" items aded in condition 2c."}

@@

```{r exp2_altsPlot_likedLoved, fig.pos = "t", fig.align='center', fig.height=2, fig.cap = "Combined counts for participant-generated alternatives for liked vs. loved in Experiment 2."}

@@

To examine the effect different alternative sets have on implicature, we needed a way of expanding alternative sets beyond entailment pairs. We addressed this issue by adopting a modified cloze task to elicit alternatives empirically. This design was inspired by Experiment 3 of @vantiel2014.

@@

```{r stimuli_exp1, fig.env = "figure*", fig.pos = "t", fig.width=6, fig.height=4, fig.align='center', set.cap.width=TRUE, num.cols.cap=2, fig.cap = "(Left) A trial from Experiment 2 (literal listener) with the target scalar \"liked.\"(Right) A trial from Experiment 3 (pragmatic listener) with the target scalar \"liked.\""}

@@

Experiment 2 was conducted to approximate literal listener semantics---$p_{L_0}(m \mid u)$ in Equation (3)---for the same five pairs of scalar items used in Experiment 1. We included three conditions: two alternatives ("entailment"), four alternatives, and five alternatives. The two alternatives condition makes a test of the hypothesis that the two members of the classic Horn (entailment) scale [@horn1972] are the only alternatives necessary to predict the strength of listeners' pragmatic inference. The four and five alternatives conditions then add successively more alternatives to test whether including a larger number of alternatives will increase model fit.\footnote{Note that alternatives in the four and five alternatives conditions were chosen on the basis of Experiment 1, which was run chronologically after the two-alternative condition; all literal listener experiments are grouped together for simplicity in reporting.} A secondary goal of Experiment 2 is to test whether the set of alternatives queried during literal semantic elicitation impacts literal semantic judgments. If it does we should see differences in these judgments for shared items between experiments.

@@

Conditions were run sequentially. In each condition we recruited 30 participants from Amazon Mechanical Turk (AMT). In the two alternatives condition, `r 30 - length(unique(exp1a$workerid))` participants were excluded for either failing to pass two training trials or because they were not native English speakers, leaving a total sample of `r length(unique(exp1a$workerid))` participants.\footnote{The majority of respondent data excluded from the two-alternatives condition was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents and made adjustments to the training trials in later conditions.} In the four alternative condition,  `r 30 - length(unique(exp1b$workerid))` participants were excluded for either failing to pass two training trials or not being native English speakers, leaving a total sample of `r length(unique(exp1b$workerid))` participants. In the five alternative condition, `r 30 - length(unique(exp1c$workerid))` participants were excluded for either failing to pass two training trials or not being native English speakers, leaving a total sample of `r length(unique(exp1c$workerid))`.

@@

The two-alternatives condition included only the two terms in the pairs from van Tiel (2014). The four-alternatives condition included the two scalar items plus the top two alternatives generated for each scalar family by participants in Experiment 1. The five-alternatives condition included the four previous items plus one more neutral item chosen from alternatives generated in Experiment 1.

@@

```{r exp1Plots, fig.env = "figure*", fig.pos = "t", fig.width=6.5, fig.height=4, fig.align='center', set.cap.width=T, num.cols.cap=2, fig.cap = "Literal listener judgments from Experiment 2. Proportion of participants indicating compatibility (answering \"yes\") is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Rows are grouped by scale and items within rows are ordered by valence. Colors indicate the specific condition (1a,b,c) with conditions including different numbers of items."}

@@

Figure \ref{fig:exp1Plots} plots estimated literal listener semantics for the three conditions. Each row shows a unique scalar family with items ordered horizontally by valence. Several trends are visible. First, in each scale, the alternatives spanned the star scale - there were scalar items that were highly compatible with both the lowest and highest numbers of stars. Second, participant compatibility judgments match intuitive literal semantics. That is, both weaker (e.g. *some* or *good*) and stronger (e.g. *all* or *excellent*) entailment items were seen as compatible with five-star ratings. This means participants' compatibility judgements reflect literal semantic intuitions, not enriched pragmatic judgments. We see clear variability between scalar families (literal semantic distributions for *memorable* are considerably different from *palatable*), but also substantial consistency across conditions (compatibility judgments for individual items are consistent across conditions).

@@

To examine this last issue of consistency across conditions (our secondary hypothesis) we fit a mixed effects model, regressing compatibility judgment on scale, number of stars and experimental condition, with subject- and word-level random effects, which was the maximal structure that converged. Results indicate no significant differences between two- and five-alternative conditions ($\beta=`r round(sumGlm2$coefficients[3], 2)`$, $z = `r round(sumGlm2$coefficients[19], 2)`$, $p = `r round(sumGlm2$coefficients[27], 2)`$) or between four- and five-alternative conditions ($\beta = `r round(sumGlm2$coefficients[2], 2)`$, $z = `r round(sumGlm2$coefficients[18], 2)`$, $p = `r round(sumGlm2$coefficients[26], 2)`$). The addition of condition as a predictor did not significantly improve model fit when compared to a model without the condition variable using ANOVA ($\chi^2(`r anovaResults$"Chi Df"[2]`) = `r round(anovaResults$Chisq[2], 2)`$, $p = `r round(anovaResults$"Pr(>Chisq)"[2], 2)`$). These findings suggest that literal semantic intuitions are not affected by the set of alternatives queried in our paradigm. This finding is important because literal semantic data generated in these three conditions are used to simulate the effects of different alternative sets on implicature generation in our model.

@@

Model fit with human judgments was significantly improved by the inclusion of alternatives beyond the entailment items (Table 1). The two-alternatives model contained only entailment items, which, under classic accounts, should be sufficient to generate implicature, but fit to data was quite poor with these items. The addition of a generic negative element produced some gains in performance, but much higher performance was found when we included four and five alternatives, with the alternatives derived empirically for the specific scale we used. An example fit for the five-alternatives model is shown in Figure \ref{fig:fiveAltsScatter}.

@@

```{r fiveAltsScatter, fig.pos = "t", fig.width=3.5, fig.height=2.25, fig.cap = "Judgments from the four-alternatives condition of Experiment 3 are plotted against model predictions using the five-alternatives data from Experiment 1. Colors show the star rating for individual judgments."}
# Correlation scatter plot
ggplot(subset(plotData, study == "Five alts"), aes(x = listener.p2, y = preds, col=stars)) + 
  geom_point() +
  xlab("model predictions") +
  ylab("human judgments") +
  stat_smooth(method = "lm") +
  theme_bw() + 
  theme(strip.text.x = element_text(size = 8),
        legend.position = "right",
        legend.margin = unit(.02,"in"),
        legend.key = element_blank(), 
        legend.key.size = unit(.1, "in"),
        axis.title.x = element_text(size = 8.5),
        axis.title.y = element_text(size = 8.5)) 
  # ggtitle("Model fit\nFive alternatives")
```

>>>>

In our current work, we use a computational framework to investigate the set of alternatives that best allow the model to predict human pragmatic judgments. We begin by presenting the computational framework (RSA) we use throughout the paper. We next describe a series of experiments designed to measure both the literal semantics of a set of scalar items (used as input to the model) and comprehenders' pragmatic judgments for these same items (used for model comparison). These experiments allow us to compare the effects of different alternative sets on our ability to model listeners' pragmatic judgments. To preview our results: we find that standard entailment alternatives do not allow us to fit participants' judgments, but that expanding the range of alternatives empirically (by asking participants to generate alternative messages) allows us to predict listeners' pragmatic judgments with high accuracy.

@@

```{r allScalesTable, fig.pos = "t", fig.width=3.5, fig.height=2.5, fig.cap = "Alternative sets used in Experiments 2 and 3 (generated in Experiment 1). Green coloring denotes classic entailment items pulled from van Tiel et al. (2014), used in the two alternative condition. Blue coloring denotes top-two alternatives added to entailment scales in the four alternative condition. Red coloring denotes ``neutral\" items aded in the five alternatives condition."}

@@

```{r exp2_altsPlot_likedLoved, fig.pos = "t", fig.align='center', fig.height=2, fig.cap = "Combined counts for participant-generated alternatives for liked vs. loved in Experiment 1."}

@@

To examine the effect different alternative sets have on implicature, we needed a way of expanding alternative sets beyond entailment pairs. We addressed this issue by adopting a modified cloze task to elicit alternatives empirically. This design was inspired by Experiment 3 of @vantiel2014b.

@@

```{r stimuli_exp1, fig.env = "figure*", fig.pos = "t", fig.width=6, fig.height=4, fig.align='center', set.cap.width=TRUE, num.cols.cap=2, fig.cap = "(Left) A trial from Experiment 2 (literal listener) with the target scalar \``liked.\"(Right) A trial from Experiment 3 (pragmatic listener) with the target scalar \``liked.\""}

@@

Experiment 2 was conducted to approximate literal listener semantics---$p_{L_0}(m \mid u)$ in Equation (3)---for the same five pairs of scalar items used in Experiment 1. We included three conditions: two alternatives ("entailment"), four alternatives, and five alternatives. The two alternatives condition makes a test of the hypothesis that the two members of the classic Horn (entailment) scale [@horn1972] are the only alternatives necessary to predict the strength of listeners' pragmatic inference. The four and five alternatives conditions then add successively more alternatives to test whether including a larger number of alternatives will increase model fit.\footnote{Note that alternatives in the four and five alternatives conditions were chosen on the basis of Experiment 1, which was run chronologically after the two alternative condition; all literal listener experiments are grouped together for simplicity in reporting.} A secondary goal of Experiment 2 is to test whether the set of alternatives queried during literal semantic elicitation impacts literal semantic judgments. If it does we should see differences in these judgments for shared items between experiments.

@@

Conditions were run sequentially. In each condition we recruited 30 participants from Amazon Mechanical Turk (AMT). In the two alternatives condition, `r 30 - length(unique(exp1a$workerid))` participants were excluded for either failing to pass two training trials or because they were not native English speakers, leaving a total sample of `r length(unique(exp1a$workerid))` participants.\footnote{The majority of respondent data excluded from the two alternatives condition was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents and made adjustments to the training trials in later conditions.} In the four alternative condition,  `r 30 - length(unique(exp1b$workerid))` participants were excluded for either failing to pass two training trials or not being native English speakers, leaving a total sample of `r length(unique(exp1b$workerid))` participants. In the five alternative condition, `r 30 - length(unique(exp1c$workerid))` participants were excluded for either failing to pass two training trials or not being native English speakers, leaving a total sample of `r length(unique(exp1c$workerid))` participants.

@@

The two alternatives condition included only the two terms in the pairs from @vantiel2014b. The four alternatives condition included the two scalar items plus the top two alternatives generated for each scalar family by participants in Experiment 1. The five alternatives condition included the four previous items plus one more neutral item chosen from alternatives generated in Experiment 1.

@@

```{r exp1Plots, fig.env = "figure*", fig.pos = "t", fig.width=6.5, fig.height=4, fig.align='center', set.cap.width=T, num.cols.cap=2, fig.cap = "Literal listener judgments from Experiment 2. Proportion of participants indicating compatibility (answering \``yes\") is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Rows are grouped by scale and items within rows are ordered by valence. Colors indicate the specific condition with conditions including different numbers of items."}

@@

Figure \ref{fig:exp1Plots} plots estimated literal listener semantics for the three conditions. Each row shows a unique scalar family with items ordered horizontally by valence. Several trends are visible. First, in each scale, the alternatives spanned the star scale - there were scalar items that were highly compatible with both the lowest and highest numbers of stars. Second, participant compatibility judgments match intuitive literal semantics. That is, both weaker (e.g. *some* or *good*) and stronger (e.g. *all* or *excellent*) entailment items were seen as compatible with five-star ratings. This means participants' compatibility judgements reflect literal semantic intuitions, not enriched pragmatic judgments. We see clear variability between scalar families (literal semantic distributions for *memorable* are considerably different from *palatable*), but also substantial consistency across conditions.

@@

To examine this last issue of consistency across conditions (our secondary hypothesis) we fit a mixed effects model, regressing compatibility judgment on scale, number of stars and experimental condition, with subject- and word-level random effects, which was the maximal structure that converged. Results indicate no significant differences between two and five alternative conditions ($\beta=`r round(sumGlm2$coefficients[3], 2)`$, $z = `r round(sumGlm2$coefficients[19], 2)`$, $p = `r round(sumGlm2$coefficients[27], 2)`$) or between four and five alternative conditions ($\beta = `r round(sumGlm2$coefficients[2], 2)`$, $z = `r round(sumGlm2$coefficients[18], 2)`$, $p = `r round(sumGlm2$coefficients[26], 2)`$). The addition of condition as a predictor did not significantly improve model fit when compared to a model without the condition variable using ANOVA ($\chi^2(`r anovaResults$"Chi Df"[2]`) = `r round(anovaResults$Chisq[2], 2)`$, $p = `r round(anovaResults$"Pr(>Chisq)"[2], 2)`$). These findings suggest that literal semantic intuitions are not affected by the set of alternatives queried in our paradigm. This finding is important because literal semantic data generated in these three conditions are used as input to our model to simulate the effects of different alternative sets on implicature generation.

@@

Model fit with human judgments was significantly improved by the inclusion of alternatives beyond the entailment items (Table 1). The two alternatives model contained only entailment items, which, under classic accounts, should be sufficient to generate implicature, but fit to data was quite poor with these items. The addition of a generic negative element produced some gains in performance, but much higher performance was found when we included four and five alternatives, with the alternatives derived empirically for the specific scale we used. An example fit for the five alternatives model is shown in Figure \ref{fig:fiveAltsScatter}.

@@

```{r fiveAltsScatter, fig.pos = "t", fig.width=3.5, fig.height=2.25, fig.cap = "Judgments for entailment items from the four alternatives condition of Experiment 3 are plotted against model predictions using the five alternatives data from Experiment 2. Colors show the star rating for individual judgments."}
# Correlation scatter plot
ggplot(subset(plotData, study == "Five alts"), aes(x = listener.p2, y = preds, col=stars)) + 
  geom_point() +
  xlab("model predictions") +
  ylab("human judgments") +
  stat_smooth(method = "lm") +
  theme_bw() + 
  theme(strip.text.x = element_text(size = 8),
        legend.position = "right",
        legend.margin = unit(.02,"in"),
        legend.key = element_blank(), 
        legend.key.size = unit(.1, "in"),
        axis.title.x = element_text(size = 8.5),
        axis.title.y = element_text(size = 8.5)) 
  # ggtitle("Model fit\nFive alternatives")
```

### 2016-02-02 Paper/Paper.tex
second to last pass fixed x- alts labeling and some wording

\begin{CodeChunk}
\begin{figure}[t]
\includegraphics{figs/allScalesTable-1} \caption[Alternative sets used in Experiments 2 and 3]{Alternative sets used in Experiments 2 and 3. Green coloring denotes classic entailment items used in conditions 2a and 3a. Blue coloring denotes top-two alternatives added to entailment scales (from Experiment 1) in conditions 2b and 3b. Red coloring denotes ``neutral" items aded in condition 2c.}\label{fig:allScalesTable}
\end{figure}
\end{CodeChunk}

@@

\caption[Combined counts for participant-generated alternatives for liked vs]{Combined counts for participant-generated alternatives for liked vs. loved in Experiment 2.}\label{fig:exp2_altsPlot_likedLoved}
\end{figure}
\end{CodeChunk}

@@

To examine the effect different alternative sets have on implicature, we
needed a way of expanding alternative sets beyond entailment pairs. We
addressed this issue by adopting a modified cloze task to elicit
alternatives empirically. This design was inspired by Experiment 3 of
{van Tiel} (2014).

@@

\caption[(Left) A trial from Experiment 2 (literal listener) with the target scalar "liked]{(Left) A trial from Experiment 2 (literal listener) with the target scalar "liked."(Right) A trial from Experiment 3 (pragmatic listener) with the target scalar "liked."}\label{fig:stimuli_exp1}
\end{figure*}
\end{CodeChunk}

@@

Experiment 2 was conducted to approximate literal listener
semantics---\(p_{L_0}(m \mid u)\) in Equation (3)---for the same five
pairs of scalar items used in Experiment 1. We included three
conditions: two alternatives (``entailment''), four alternatives, and
five alternatives. The two alternatives condition makes a test of the
hypothesis that the two members of the classic Horn (entailment) scale
(Horn, 1972) are the only alternatives necessary to predict the strength
of listeners' pragmatic inference. The four and five alternatives
conditions then add successively more alternatives to test whether
including a larger number of alternatives will increase model
fit.\footnote{Note that alternatives in the four and five alternatives conditions were chosen on the basis of Experiment 1, which was run chronologically after the two-alternative condition; all literal listener experiments are grouped together for simplicity in reporting.}
A secondary goal of Experiment 2 is to test whether the set of
alternatives queried during literal semantic elicitation impacts literal
semantic judgments. If it does we should see differences in these
judgments for shared items between experiments.

@@

Conditions were run sequentially. In each condition we recruited 30
participants from Amazon Mechanical Turk (AMT). In the two alternatives
condition, 16 participants were excluded for either failing to pass two
training trials or because they were not native English speakers,
leaving a total sample of 14
participants.\footnote{The majority of respondent data excluded from the two-alternatives condition was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents and made adjustments to the training trials in later conditions.}
In the four alternative condition, 7 participants were excluded for
either failing to pass two training trials or not being native English
speakers, leaving a total sample of 23 participants. In the five
alternative condition, 3 participants were excluded for either failing
to pass two training trials or not being native English speakers,
leaving a total sample of 27.

@@

The two-alternatives condition included only the two terms in the pairs
from van Tiel (2014). The four-alternatives condition included the two
scalar items plus the top two alternatives generated for each scalar
family by participants in Experiment 1. The five-alternatives condition
included the four previous items plus one more neutral item chosen from
alternatives generated in Experiment 1.

@@

\caption[Literal listener judgments from Experiment 2]{Literal listener judgments from Experiment 2. Proportion of participants indicating compatibility (answering "yes") is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Rows are grouped by scale and items within rows are ordered by valence. Colors indicate the specific condition (1a,b,c) with conditions including different numbers of items.}\label{fig:exp1Plots}
\end{figure*}
\end{CodeChunk}

Figure \ref{fig:exp1Plots} plots estimated literal listener semantics
for the three conditions. Each row shows a unique scalar family with
items ordered horizontally by valence. Several trends are visible.
First, in each scale, the alternatives spanned the star scale - there
were scalar items that were highly compatible with both the lowest and
highest numbers of stars. Second, participant compatibility judgments
match intuitive literal semantics. That is, both weaker (e.g.
\emph{some} or \emph{good}) and stronger (e.g. \emph{all} or
\emph{excellent}) entailment items were seen as compatible with
five-star ratings. This means participants' compatibility judgements
reflect literal semantic intuitions, not enriched pragmatic judgments.
We see clear variability between scalar families (literal semantic
distributions for \emph{memorable} are considerably different from
\emph{palatable}), but also substantial consistency across conditions
(compatibility judgments for individual items are consistent across
conditions).

To examine this last issue of consistency across conditions (our
secondary hypothesis) we fit a mixed effects model, regressing
compatibility judgment on scale, number of stars and experimental
condition, with subject- and word-level random effects, which was the
maximal structure that converged. Results indicate no significant
differences between two- and five-alternative conditions
(\(\beta=-0.05\), \(z = -0.53\), \(p = 0.59\)) or between four- and
five-alternative conditions (\(\beta = -0.04\), \(z = -0.52\),
\(p = 0.6\)). The addition of condition as a predictor did not
significantly improve model fit when compared to a model without the
condition variable using ANOVA (\(\chi^2(2) = 0.43\), \(p = 0.81\)).
These findings suggest that literal semantic intuitions are not affected
by the set of alternatives queried in our paradigm. This finding is
important because literal semantic data generated in these three
conditions are used to simulate the effects of different alternative
sets on implicature generation in our model.

@@

\begin{CodeChunk}
\begin{figure}[t]
\includegraphics{figs/fiveAltsScatter-1} \caption[Judgments from the four-alternatives condition of Experiment 3 are plotted against model predictions using the five-alternatives data from Experiment 1]{Judgments from the four-alternatives condition of Experiment 3 are plotted against model predictions using the five-alternatives data from Experiment 1. Colors show the star rating for individual judgments.}\label{fig:fiveAltsScatter}
\end{figure}
\end{CodeChunk}

>>>>

\begin{CodeChunk}
\begin{figure}[t]
\includegraphics{figs/allScalesTable-1} \caption[Alternative sets used in Experiments 2 and 3 (generated in Experiment 1)]{Alternative sets used in Experiments 2 and 3 (generated in Experiment 1). Green coloring denotes classic entailment items pulled from van Tiel et al. (2014), used in the two alternative condition. Blue coloring denotes top-two alternatives added to entailment scales in the four alternative condition. Red coloring denotes ``neutral" items aded in the five alternatives condition.}\label{fig:allScalesTable}
\end{figure}
\end{CodeChunk}

@@

\caption[Combined counts for participant-generated alternatives for liked vs]{Combined counts for participant-generated alternatives for liked vs. loved in Experiment 1.}\label{fig:exp2_altsPlot_likedLoved}
\end{figure}
\end{CodeChunk}

@@

To examine the effect different alternative sets have on implicature, we
needed a way of expanding alternative sets beyond entailment pairs. We
addressed this issue by adopting a modified cloze task to elicit
alternatives empirically. This design was inspired by Experiment 3 of
{van Tiel} et al. (2014).

@@

\caption[(Left) A trial from Experiment 2 (literal listener) with the target scalar ``liked]{(Left) A trial from Experiment 2 (literal listener) with the target scalar ``liked."(Right) A trial from Experiment 3 (pragmatic listener) with the target scalar ``liked."}\label{fig:stimuli_exp1}
\end{figure*}
\end{CodeChunk}

@@

Experiment 2 was conducted to approximate literal listener
semantics---\(p_{L_0}(m \mid u)\) in Equation (3)---for the same five
pairs of scalar items used in Experiment 1. We included three
conditions: two alternatives (``entailment''), four alternatives, and
five alternatives. The two alternatives condition makes a test of the
hypothesis that the two members of the classic Horn (entailment) scale
(Horn, 1972) are the only alternatives necessary to predict the strength
of listeners' pragmatic inference. The four and five alternatives
conditions then add successively more alternatives to test whether
including a larger number of alternatives will increase model
fit.\footnote{Note that alternatives in the four and five alternatives conditions were chosen on the basis of Experiment 1, which was run chronologically after the two alternative condition; all literal listener experiments are grouped together for simplicity in reporting.}
A secondary goal of Experiment 2 is to test whether the set of
alternatives queried during literal semantic elicitation impacts literal
semantic judgments. If it does we should see differences in these
judgments for shared items between experiments.

@@

Conditions were run sequentially. In each condition we recruited 30
participants from Amazon Mechanical Turk (AMT). In the two alternatives
condition, 16 participants were excluded for either failing to pass two
training trials or because they were not native English speakers,
leaving a total sample of 14
participants.\footnote{The majority of respondent data excluded from the two alternatives condition was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents and made adjustments to the training trials in later conditions.}
In the four alternative condition, 7 participants were excluded for
either failing to pass two training trials or not being native English
speakers, leaving a total sample of 23 participants. In the five
alternative condition, 3 participants were excluded for either failing
to pass two training trials or not being native English speakers,
leaving a total sample of 27 participants.

@@

The two alternatives condition included only the two terms in the pairs
from {van Tiel} et al. (2014). The four alternatives condition included
the two scalar items plus the top two alternatives generated for each
scalar family by participants in Experiment 1. The five alternatives
condition included the four previous items plus one more neutral item
chosen from alternatives generated in Experiment 1.

@@

\caption[Literal listener judgments from Experiment 2]{Literal listener judgments from Experiment 2. Proportion of participants indicating compatibility (answering ``yes") is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Rows are grouped by scale and items within rows are ordered by valence. Colors indicate the specific condition with conditions including different numbers of items.}\label{fig:exp1Plots}
\end{figure*}
\end{CodeChunk}

Figure \ref{fig:exp1Plots} plots estimated literal listener semantics
for the three conditions. Each row shows a unique scalar family with
items ordered horizontally by valence. Several trends are visible.
First, in each scale, the alternatives spanned the star scale - there
were scalar items that were highly compatible with both the lowest and
highest numbers of stars. Second, participant compatibility judgments
match intuitive literal semantics. That is, both weaker (e.g.
\emph{some} or \emph{good}) and stronger (e.g. \emph{all} or
\emph{excellent}) entailment items were seen as compatible with
five-star ratings. This means participants' compatibility judgements
reflect literal semantic intuitions, not enriched pragmatic judgments.
We see clear variability between scalar families (literal semantic
distributions for \emph{memorable} are considerably different from
\emph{palatable}), but also substantial consistency across conditions.

To examine this last issue of consistency across conditions (our
secondary hypothesis) we fit a mixed effects model, regressing
compatibility judgment on scale, number of stars and experimental
condition, with subject- and word-level random effects, which was the
maximal structure that converged. Results indicate no significant
differences between two and five alternative conditions
(\(\beta=-0.05\), \(z = -0.53\), \(p = 0.59\)) or between four and five
alternative conditions (\(\beta = -0.04\), \(z = -0.52\), \(p = 0.6\)).
The addition of condition as a predictor did not significantly improve
model fit when compared to a model without the condition variable using
ANOVA (\(\chi^2(2) = 0.43\), \(p = 0.81\)). These findings suggest that
literal semantic intuitions are not affected by the set of alternatives
queried in our paradigm. This finding is important because literal
semantic data generated in these three conditions are used as input to
our model to simulate the effects of different alternative sets on
implicature generation.

@@

\begin{CodeChunk}
\begin{figure}[t]
\includegraphics{figs/fiveAltsScatter-1} \caption[Judgments for entailment items from the four alternatives condition of Experiment 3 are plotted against model predictions using the five alternatives data from Experiment 2]{Judgments for entailment items from the four alternatives condition of Experiment 3 are plotted against model predictions using the five alternatives data from Experiment 2. Colors show the star rating for individual judgments.}\label{fig:fiveAltsScatter}
\end{figure}
\end{CodeChunk}
